---
title: "Hierarchical Clustering"
output:
  html_document: default
  pdf_document: default
---



```{r include=FALSE}
#Replace the below command to point to your working directory
setwd ("/Users/kumar/Dropbox/_Projects/GreatLearning/GL-scripts/")
```


We will work with this simple Customer Spend Data:
```{r}
custSpendData <- read.csv("D:/Prasanna/Personal/Learning/Other Reads/GL/Course Material/Data Mining/DataSets/Cust_Spend_Data.csv", header=TRUE)
print(custSpendData)
```


The dist function computes the distance between row vectors of a given matrix.  Several distance functions can be used. 
```{r}
distMatrix = dist(x=custSpendData[,3:7], method = "euclidean") 
#distMatrix = dist(x=custSpendData[,3:7], method = "minkowski", p=2)   #equivalent to above line
print(distMatrix, digits = 3)
```

Notice that the Avg_Mnthly_Spend column dominates the distance since its magnitude is much higher than the other columns. Since it there is no reason to over-weight one column than another, we scale all the columns to mean zero and unit standard deviation.

```{r}
## scale function standardizes the values
custSpendData.Scaled = scale(custSpendData[,3:7])
print(custSpendData.Scaled)
```


Checking the scaling
```{r}
apply(custSpendData.Scaled,2,mean)
apply(custSpendData.Scaled,2,sd)
```

and the scaled distance matrix
```{r}
# Compute distance matrix again with scaled data
distMatrix.Scaled = dist(x=custSpendData.Scaled, method = "euclidean") 
print(distMatrix.Scaled, digits = 3)
```


Hierarchical Clustring with hclust
```{r}
cluster <- hclust(distMatrix.Scaled, method = "average")
plot(cluster, labels = as.character(custSpendData[,2]))
```


The cluster combining heights
```{r}
cluster$height
```


Now we can choose a K, lets say 3, and draw boxes around the 3 clusters
```{r}
# Plot rectagles defining the clusters for any given level of K 
plot(cluster, labels = as.character(custSpendData[,2]))
rect.hclust(cluster, k=2, border="red")
```


Lets add the cluter number field back to the original dataset

```{r}
custSpendData$Cluster <- cutree(cluster, k=3)
print(custSpendData)
```

Finally lets aggregate the three types of customers by averaging columns 3 to 7

```{r}
## Aggregate columns 3:7 for each cluster by their means
custProfile = aggregate(custSpendData[,-c(1,2, 8)],list(custSpendData$Cluster),FUN="mean")
print(custProfile)
```

